Treatment heterogeneity in comparative effectiveness of teriparatide vs bisphosphonates: multi-database cohort study

### 2021-02-27 submission/manuscript.rmd
add figure to plot and format numbers

---
title: "Treatment heterogeneity in comparative effectiveness of teriparatide vs bisphosphonates: multi-database cohort study"
output:
    bookdown::pdf_document2: default
    bookdown::word_document2:
        reference_docx: reference.docx
geometry: margin=1.0in
toc: false
font-size: 11pt
header-includes:
  - \usepackage{helvet}
  - \renewcommand*\familydefault{\sfdefault}
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage[left]{lineno}
  - \linenumbers
editor_options: 
  chunk_output_type: console
---

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(knitr)
options(scipen = 1, digits = 2)
knit_hooks$set(
  inline = function(x) {
    prettyNum(
      x, 
      big.mark = ",",
      decimal.mark = ".",
      preserve.width = "individual"
    )
  }
)
```

\newpage
# Introduction
Osteoporosis is a common condition characterised by decreased bone density and associated with increased risk for fragility fractures, which affects almost 30% of women aged $\geq50$ years (#Wright et al. 2017). It was estimated that in 2005, over 2 million incident fractures occurred in the United States, with hip fractures accounting for more than 70% of the costs. By 2025, annual fractures and costs could rise by almost 50% (#Cole et al. 2008).

@@

##     Statistical analyses
We performed two sets of analyses. First, we derived overall treatment effect estimates of teriparatide compared to oral bisphosphonates regarding the three outcomes of interest. To account for potential measured confounding we developed separate propensity score models within each database using the same predefined set of measured covariates. Covariates were included in the propensity score model using regularized logistic regression (##REFERENCE Tian-Schuemie simulation). We estimated treatment effects using Cox proportional hazards models with treatment as the sole covariate fitted in the 1-4 propensity score-matched subset of the considered patient population. Results from different databases were summarized using random effects meta-analysis (##REFERECNE).

@@

```{r, echo=FALSE}
metaCalibrateOverall <- readRDS(
  here::here("data/processed/metaCalibrateOverall.rds")
)
```

The random effects meta-analysis of teriparatide compared to oral bisphosphonates across the 3 considered databases gave a calibrated overall hazard ratio of `r metaCalibrateOverall %>% filter(outcome == 101) %>% pull(hr)` (`r metaCalibrateOverall %>% filter(outcome == 101) %>% pull(lower)` to `r metaCalibrateOverall %>% filter(outcome == 101) %>% pull(upper)`; 95% CI) for hip fracture, `r metaCalibrateOverall %>% filter(outcome == 102) %>% pull(hr)` (`r metaCalibrateOverall %>% filter(outcome == 102) %>% pull(lower)` to `r metaCalibrateOverall %>% filter(outcome == 102) %>% pull(upper)`; 95%CI) for major osteoporotic fracture and `r metaCalibrateOverall %>% filter(outcome == 103) %>% pull(hr)` (`r metaCalibrateOverall %>% filter(outcome == 103) %>% pull(lower)` to `r metaCalibrateOverall %>% filter(outcome == 103) %>% pull(upper)`; 95%CI) for vertebral fracture (##Figures 1-3).

```{r, echo=FALSE} 
metaCalibrateRiskStratified <- readRDS(
  here::here("data/processed/metaCalibrateRiskStratified.rds")
)

calibrateRiskStratified <- readRDS(
  here::here("data/processed/calibrateRiskStratified.rds")
) %>%
  mutate(
    hr = exp(logRr),
    lower = exp(logLb95Rr),
    upper = exp(logUb95Rr)
  )

hipFractureAbsolute <- readRDS(
  here::here("data/processed/hipFractureAbsolute.rds")
) %>%
  filter(riskStratum == "Q4") %>%
  mutate(
    estimate = 100 * estimate,
    lower    = 100 * lower,
    upper    = 100 * upper
  )

@@

>>>>

---
title: "Treatment heterogeneity in comparative effectiveness of teriparatide vs bisphosphonates: multi-database cohort study"
output:
    bookdown::pdf_document2: default
    bookdown::word_document2:
        reference_docx: reference.docx
geometry: margin=1.0in
toc: false
font-size: 11pt
header-includes:
  - \renewcommand*\familydefault{\sfdefault}
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage[left]{lineno}
editor_options: 
  chunk_output_type: console
---

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(pander)
library(knitr)

d <- function(x, decimals = 2) {
  sprintf(paste0("%1.", decimals, "f"), x) 
}

knit_hooks$set(
  inline = function(x) {
    prettyNum(
      x,
      big.mark = ",",
      decimal.mark = ".",
      preserve.width = "individual"
    )
  }
)
```

\vspace{35mm}

Alexandros Rekkas${^1}$, Annika M. J&ouml;dicke${^2}$, David van Klaveren$^3$, Daniel Prieto-Alhambra${^{1,2}}$, Peter R. Rijnbeek${^1}$

\vspace{40mm}

$^1$ Department of Medical Informatics, Erasmus University Medical Center, Rotterdam, Netherlands

$^2$ Pharmaco- and Device Epidemiology, Centre for Statistics in Medicine, Nuffield Department of Orthopaedics, Rheumatology and Musculoskeletal Sciences, University of Oxford, Oxford, UK

$^3$ Department of Public Health, Erasmus University Medical Center, Rotterdam, Netherlands

\newpage
\linenumbers
# Introduction
Osteoporosis is a common condition characterised by decreased bone density and associated with increased risk for fragility fractures, which affects almost 30% of women aged $\geq50$ years (#Wright et al. 2017). It was estimated that in 2005, over 2 million incident fractures occurred in the United States, with hip fractures accounting for more than 70% of the costs. By 2025, annual fractures and costs could rise by almost 50% (#Cole et al. 2008).

@@

##     Statistical analyses
We carried out two sets of analyses. First, we derived overall treatment effect estimates of teriparatide compared to oral bisphosphonates regarding the three outcomes of interest. To account for potential measured confounding we developed separate propensity score models within each database using the same predefined set of measured covariates. Covariates were included in the propensity score model using regularized logistic regression (##REFERENCE Tian-Schuemie simulation). We estimated treatment effects using Cox proportional hazards models with treatment as the sole covariate fitted in the 1-4 propensity score-matched subset of the considered patient population. Results from different databases were summarized using random effects meta-analysis (##REFERECNE).

@@

```{r, echo=FALSE}
metaCalibrateOverall <- readRDS(
  here::here("data/processed/metaCalibrateOverall.rds")
) %>%
  mutate(
    hr = d(hr),
    lower = d(lower),
    upper = d(upper)
  )
```

The random effects meta-analysis of teriparatide compared to oral bisphosphonates across the 3 considered databases gave a calibrated overall hazard ratio of `r metaCalibrateOverall %>% filter(outcome == 101) %>% pull(hr)` (`r metaCalibrateOverall %>% filter(outcome == 101) %>% pull(lower)` to `r metaCalibrateOverall %>% filter(outcome == 101) %>% pull(upper)`; 95% CI) for hip fracture, `r metaCalibrateOverall %>% filter(outcome == 102) %>% pull(hr)` (`r metaCalibrateOverall %>% filter(outcome == 102) %>% pull(lower)` to `r metaCalibrateOverall %>% filter(outcome == 102) %>% pull(upper)`; 95%CI) for major osteoporotic fracture and `r metaCalibrateOverall %>% filter(outcome == 103) %>% pull(hr)` (`r metaCalibrateOverall %>% filter(outcome == 103) %>% pull(lower)` to `r metaCalibrateOverall %>% filter(outcome == 103) %>% pull(upper)`; 95%CI) for vertebral fracture (Figure \@ref(fig:plotMeta)).

```{r, echo=FALSE} 
metaCalibrateRiskStratified <- readRDS(
  here::here("data/processed/metaCalibrateRiskStratified.rds")
) %>%
  mutate(
    hr = d(hr),
    lower = d(lower),
    upper = d(upper)
  )

calibrateRiskStratified <- readRDS(
  here::here("data/processed/calibrateRiskStratified.rds")
) %>%
  mutate(
    hr = d(exp(logRr)),
    lower = d(exp(logLb95Rr)),
    upper = d(exp(logUb95Rr))
  )

hipFractureAbsolute <- readRDS(
  here::here("data/processed/hipFractureAbsolute.rds")
) %>%
  filter(riskStratum == "Q4") %>%
  mutate(
    estimate = d(100 * estimate),
    lower    = d(100 * lower),
    upper    = d(100 * upper)
  )

@@

\newpage
```{r plotMeta, echo=FALSE, fig.cap="This is a caption", out.width = '100%'}
knitr::include_graphics(here::here("figures/plotMeta.pdf"))
```
